---
title: "Package Installation"
author: "nabil Abd..."
date: "May 19, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

To perform processing with the hybrid optimization, using the `hybridSA` package, 
there are a number of packages you should have installed for your `R` environment. 

Running the following commands should install most of these packages on your 
machine: 

```{r}
pkgs <- c("ggplot2", "knitr", "devtools", "rmarkdown", "spacetime", "mapproj", 
          "purrr", "pryr", "readr", "beepr", "lubridate", "magrittr", 
          "assertthat", "nloptr", "dplyr", "tidyr", "automap")

install.packages(pkgs)
devtools::install_github("hadley/dplyr") # fixes bugs in version 0.4.3
```

Additionally, you will also need to install the `rgdal` and `gstat` packages. Depending on 
the operating system you're using, this might require different things. Below are some 
possible steps you might try, but since software is constantly being updated, you 
might need to look elsewhere to address the specific problems you might be having. 

## Installing `rgdal`

For the `rgdal` package to work properly, you should also have `gdal` installed on your machine. 

* If you have a Mac, and  have `homebrew`, then you can just run `brew install gdal`. If you have a 
Mac but don't have `homebrew`, why not? First get `homebrew`, then proceed with this installation process.

* If you have a Windows or a Linux, please see online resources for how to install `gdal` onto your machine.

Once you have `gdal` installed, then try running `install.packages` as above. If that doesn't work and yields some error, then try: 

```{r}
install.packages("rgdal", type = "source")
```

If you get a message saying that the `rgeos` library needs to be installed, but that fails, also try installing `geos` by 

```
brew install geos
```

from the command line.

CHECK revdeps OF GSTAT AND RGDAL


## Installing `gstat`

First try to install `gstat` via `install.packages` as above (first without `type = "source"`). 


